---
title: "hw_4"
author: "Colby S."
date: '2022-06-08'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
```

HW 4
ECP – 3004 (Summer 2022)
Due: 06/16/2022
Berk & DeMarzo: Chapter 6

Problems 1-4, 10, 13, 17-19, 28


### 8-1. done
A 30-year bond with a face value of 1000 has a coupon rate of 5.5%, with semiannual payments.
a. What is the coupon payment for this bond?
$$ CPN = \frac{\text{Coupon Rate}*\text{Face Value}}{\text{Number of Coupons per Year}}$$
b. Draw the cash flows for the bond on a timeline.


### 8-2. done 
Assume that a bond will make 20 dollar payments every six months for 20 periods, and when it expires it will pay out an additional 1000 dollars
```{r}
CPN <- 20 #dollars
face_value <- 1000 #dollars
coupons_per_year <- 2 #semiannual
maturity_periods <- 20 #maturity in periods
```

a. What is the maturity of the bond (in years)?

>the maturity, or term, of the bond is the remaining time until the bond reaches maturity. In this case we convert 20 six month periods into years.

```{r}
6*20/12 #years
```

b. What is the coupon rate (in percent)?
$$CPN = \frac{\text{Coupon Rate} * \text{Face Value}}{\text{Number of Coupons per Year}}$$
```{r}
coupon_rate <- CPN *coupons_per_year / face_value
coupon_rate
```

c. What is the face value?
```{r}
face_value #given in problem
```


### 8-3. done
The following table summarizes prices of various default-free, zero-coupon bonds (expressed as
a percentage of face value):
```{r}
maturity <- 1:5
price <- c(95.51, 91.05, 86.38, 81.65, 76.51) #per 100 dollar face value
rbind(maturity, price)
```


a. Compute the yield to maturity for each bond.
>because the bonds are zero-coupon bonds, the present value of the bond is the discounted future value of the bond. 

$$PV = \frac{FV}{\text{yeild to maturity (YTM)}^n}$$
```{r}
YTM <- round((100/price)^(1/maturity) - 1, digits = 3)
YTM
```

b. Plot the zero-coupon yield curve (for the first five years).
```{r}
data.frame(maturity, YTM) |>
ggplot(aes(x = maturity, y = YTM)) +
  geom_line()
```

c. Is the yield curve upward sloping, downward sloping, or flat?

>upword sloping


### 8-4. done
Suppose the current zero-coupon yield curve for risk-free bonds is as follows:
```{r}
maturity <- 1:5
YTM <- c(5, 5.5, 5.75, 5.95, 6.05) * 0.01 
rbind(maturity, YTM)
```
$$CPN = \frac{\text{Coupon Rate * Face Value}}{\text{Coupons per Year}}$$
$$PV = FV/(1+r)^n$$
a. What is the price per 100 face value of a two-year, zero-coupon, risk-free bond?
b. What is the price per 100 face value of a four-year, zero-coupon, risk-free bond?
c. What is the risk-free interest rate for a five-year maturity?

```{r}
n <- c(2, 4, 5)
solutions <- 100/(1+YTM[n])^n 
names(solutions) <- c("a", "b", "c")
solutions
```

### 8-10. done
Suppose a seven-year, 1000 bond with an 8% coupon rate and semiannual coupons is trading
with a yield to maturity of 6.75%.
```{r}
term <- 7 #years
face_value <- 1000 #dollars
coupon_rate <- 0.08
periods_per_year <- 2 #semiannual
YTM <- 0.0675
```


a. Is this bond currently trading at a discount, at par, or at a premium? Explain.
> a bond traving above par will have a price > face value. This occurs when the coupon rate > YTM
> a bond trading at par will have a price = face value. This occurs when the coupon rate = YTM
> a bond trading below par will have a price < face value. This occurs when teh coupon rate is < YTM

>this bond is trading at a premium because the coupon rate > YTM

b. If the yield to maturity of the bond rises to 7% (APR with semiannual compounding), what
price will the bond trade for?
$$\text{(6.5) Yeild to Maturity of a Coupon Bond}$$
$$price = \frac{CPN}{y}(1-\frac{1}{(1+y)^n}) + \frac{\text{Face Value}}{(1+y)^n}$$
```{r}
YTM <- 0.07 #APR
y <- YTM/periods_per_year
n <- term * periods_per_year # compounding periods left on bond
CPN <- coupon_rate*face_value / periods_per_year

CPN/y*(1-1/(1+y)^n) + face_value/(1+y)^n
```


### 8-13. done
Consider the following  \$100 annual bonds:
```{r, echo = FALSE}
face_value <- 100
bond <- LETTERS[1:4]
coupon_rate <- c(0, 0, 4, 8) * 0.01
maturity <- c(15, 10, 15, 10) #years
bond_df <- data.frame(bond, coupon_rate, maturity, face_value) |>
  mutate(CPN = face_value*coupon_rate)
bond_df
```

a. What is the percentage change in the price of each bond if its yield to maturity falls from
6% to 5%?
$$\text{(6.5) Yeild to Maturity of a Coupon Bond}$$
$$price = \frac{CPN}{y}(1-\frac{1}{(1+y)^n}) + \frac{\text{Face Value}}{(1+y)^n}$$

$$\text{percentage change} = \frac{\Delta price}{price_\text{original}}*100$$
```{r}
term <- maturity #the bonds are new, so the term is the full term of the bond (maturity)

pct_change_df <- bond_df |>
  mutate(price_5 = CPN/0.05*(1-1/(1+0.05)^term) + face_value/(1+0.05)^term, #price with YTM = 5%
         price_6 = CPN/0.06*(1-1/(1+0.06)^term) + face_value/(1+0.06)^term, #price with YTM = 6%
         percent_change = round((price_5 - price_6)/price_6*100, 2)) |>
  select(bond, percent_change)

# there is a more concise way to program this
YTM <- c(0.06, 0.05)
bond_df |>
  mutate(percent_change = diff(CPN/YTM*(1-1/(1+YTM)^term) + face_value/(1+YTM)^term), #price with YTM = 5%
         price_6 = CPN/0.06*(1-1/(1+0.06)^term) + face_value/(1+0.06)^term) #price with YTM = 6%

```

b. Which of the bonds A–D is most sensitive to a 1% drop in interest rates from 6% to 5% and
why? Which bond is least sensitive? Provide an intuitive explanation for your answer.
```{r}
pct_change_df[which.max(pct_change_df$percent_change), ]

pct_change_df[which.min(pct_change_df$percent_change), ]
```

### 8-17. 
What is the price today of a two-year, default-free security with a face value of \$1000 and an
annual coupon rate of 6%? Does this bond trade at a discount, at par, or at a premium?
```{r}
term <- 2 #years
face_val <- 1000 #dollars
coupon_rate <- 0.06 #6%
```
>A bond trades at a premium when the price > face value. This occurs when the coupon rate > YTM.

```{r}

```


### 8-18. 
What is the price of a five-year, zero-coupon, default-free security with a face value of 1000?
```{r}

```

### 8-19. 
What is the price of a three-year, default-free security with a face value of 1000 and an annual
coupon rate of 4%? What is the yield to maturity for this bond?
The following table summarizes the yields to maturity on several one-year, zero-coupon securities:
a. What is the price (expressed as a percentage of the face value) of a one-year, zero-coupon
corporate bond with a AAA rating?
b. What is the credit spread on AAA-rated corporate bonds?
c. What is the credit spread on B-rated corporate bonds?
d. How does the credit spread change with the bond rating? Why?

### 8-28. 
Grummon Corporation has issued zero-coupon corporate bonds with a five-year maturity. Investors believe there is a 25% chance that Grummon will default on these bonds. If Grummon does default, investors expect to receive only 65 cents per dollar they are owed. If investors require a 6% expected return on their investment in these bonds, what will be the
price and yield to maturity on these bonds?