---
title: "hw_1"
author: "Colby S."
date: '2022-06-18'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

ECP 3004
HW 1
Due Thursday 05/26/22

### 1. 
Casper Milktoast has $200,000 available to support consumption in periods 0 (now)
and 1 (next year). He wants to consume exactly the same amount in each period. The
interest rate is 8 percent. There is no risk. How much should he invest, and how much can
he consume in each period?

$$y = C_1 + \frac{c_2}{1+r}$$
$$C_1 = C_2$$
$$y = C_1 + \frac{c_1}{1+r}$$
$$y = C_1(1+1/(1+r))$$
$$C_1 = \frac{y}{(1+1/(1+r))}$$
```{r}
y <- 200000 #wealth
r <- 0.08 #interest rate

c_1 <- y / (1+1/(1+r))
c_1
c_2 <- c_1/(1+r)
c_2
```

### 2. 
An isolated society contains two groups of consumers (the same number of each kind).
The consumers all have identical preferences with respect to present and future
consumption, but they have different endowments. Group A has an endowment of (240,
160) for present and future. Group B has and endowment of (320, 440).

Their preferences can be represented by the utility function

$$U(C_0,C_1)=(C_0*C_1)^{0.5}$$

These preferences have a marginal rate of substitution equal to

$$MRS=-\frac{C_1}{C_0}$$

a. 

What is the budget constraint for members of the A group if the interest rate at which
they can borrow or lend is r?

$$ C_0+\frac{C_1}{1+r} = Y_0+\frac{Y_1}{1+r} = W$$

b. 

At an interest rate of r what is the value of a member of group A's wealth?

$$wealth_A = 240 + 160/(1+r)$$

c. 

What conditions characterize the optimal consumption point for the consumers if there
are exchange opportunities between the present and future at a real rate of interest r?

* The optimal consumption point will lie on the budget constraint

|           justification: outside the budget constraint is unaffordable and inside the budget constraint violates the more is better property

$$\text{The Budget Constraint}$$
$$ C_0+\frac{C_1}{1+r} = Y_0+\frac{Y_1}{1+r} = W$$

|           We must recognize that the budget constraint is a function with x axis $C_0$ and y axis $C_1$. 

|           We can solve for the slope of the budget constraint by putting the budget constraint in point-slope form.

$$W =  C_0+\frac{C_1}{1+r}$$
$$C_1 = -(1+r)C_0 + (1+r)W$$


* The optimal consumption point will have a slope equal to the marginal rate of substitution

|         justification: The marginal rate of substitution is the slope of the budget constraint. If the optimal consumption point lies on the budget constraint, it will inherit its slope.


d. 

Show that the consumption function for present period for members of both groups is $C_0=0.5W$ where W is the consumer's wealth. Write an expression for the saving function for each group in a form that shows its dependence on the rate of interest.

$$\text{show that C_0=0.5W}$$
$$W = C_0+\frac{C_1}{1+r}$$
from the earlier problem we know,
$$MRS = -\frac{C_1}{C_0} = -(1+r)$$
$$C1 = C_0*(1+r)$$
substituting into the first equation we get
$$W = C_0+\frac{C_0*(1+r)}{1+r}$$
$$W = 2C_0$$
$$C_0 = 0.5W$$

$$\text{Savings Function}$$
$$Savings = (Y_0-C_0) + \frac{(Y_1-C_1)}{1+r}$$
$$Savings = Y_0-C_0+\frac{Y_1}{1+r}-C_0$$
$$Savings = Y_0-2C_0+\frac{Y_1}{1+r}$$

e. What condition must hold if r is to be an equilibrium interest rate for this society,
assuming that they can exchange only among themselves. What is the equilibrium real
interest rate?


### 3. 
Suppose an individual has the following utility function

$$U(C_0,C_1) = C_0^\alpha C_1^{1-\alpha}$$
$$0<\alpha<1$$
Let the endowments be Y0 and Y1 and let r denote the interest rate.
a. Use calculus to show that

$$MRS = -\frac{U_0}{U_1} = -\frac{\alpha}{1-\alpha}\frac{C_1}{C_0}$$
$$C_0^* = \alpha W_0$$
$$C_1^* = (1-\alpha)(1+r)W_0$$
